### 2009-04-20 paper.tex
moved Ortheus placeholder comment
further fixed up null cycle section

We refer to empty paths and empty bifurcations as {\bf null cycles}.
They occur (in simpler form) in Hidden Markov Models (which are special cases of SCFGs)
and independent-sites models (which can be viewed as special cases of HMMs).
In all such models, null cycles represent events which are allowed by the model, but could never be observed.
They can be accounted for by exact marginalization.
For example, eliminating missing ``empty'' columns from an independent-sites model of sequence involves a straightforward application of a correction factor
to account for the proportion of columns that are unobserved \cite{RivasEddy2008}.
The slightly more complicated situation of missing emissions in a HMM can be dealt with by
summing over all null paths, yielding a geometric series that is solvable by matrix inversion
\cite{Holmes2003,Holmes2007,Lunter2007}; the ``wing retraction'' algorithms used in HMMER can be viewed as a special case of matrix inversion by iterated elimination \cite{hmmer}.

@@

% to do: discuss the possibility of progressive reconstruction a la Ortheus

@@

>>>>

We refer to empty paths and empty bifurcations as {\bf null cycles}.
They occur (in simpler form) in Hidden Markov Models (which are special cases of SCFGs)
and independent-sites models (which can be viewed as special cases of HMMs).
In all such models, null cycles represent histories that are valid according to the model, but cannot be resolved by direct observation.
Such null events can be real (e.g. ancestral residues that have been deleted in all extant lineages)
or they can be artefactual (e.g. transitions between placeholder null states of an HMM).
Often, one can account for the likelihood of such events by exact marginalization.
For example, eliminating missing ``empty'' columns from an independent-sites model of sequence involves a straightforward application of a correction factor
to account for the proportion of columns that are unobserved \cite{RivasEddy2008}.
The slightly more complicated situation of missing emissions in a HMM can be dealt with by
summing over all null paths, yielding a geometric series that is solvable by matrix inversion
\cite{Holmes2003,Holmes2007,Lunter2007}; the ``wing retraction'' algorithms used in HMMER can be viewed as a special case of matrix inversion by iterated elimination \cite{hmmer}.

@@

@@

% to do: discuss the possibility of progressive reconstruction a la Ortheus
% "Another heuristic that could be tried..."
